---
title: "HW6"
author: "My An Huynh"
date: "2024-11-19"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(modelr)
library(mgcv)
library(SemiPar)
library(glmnet)
library(GGally)
```

## Problem 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}

```

## Problem 2
```{r data wrangling}
homicide_df =
  read_csv("data/homicide-data.csv") |> 
  mutate(
    city_state = paste(city, state, sep = ", "),
    solved_bin = ifelse(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age)
  ) |>
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")),
    victim_race %in% c("White", "Black")
  ) |> 
  select(-city, -state)


```

Glm for Baltimore, MD
```{r}
baltimore = 
  homicide_df |> 
  filter(
    city_state == "Baltimore, MD"
  ) 

baltimore_glm = 
  glm(solved_bin ~ victim_age + victim_race + victim_sex, data = baltimore, family = binomial()) |> 
  broom::tidy() |> 
  filter(
    term == "victim_sexMale"
  ) |> 
  mutate(
    odds_ratio = exp(estimate),
    conf_upper_95 = exp(estimate + 1.96 * std.error),
    conf_lower_95 = exp(estimate - 1.96 * std.error)
  )
```

Do this for every city 

```{r}
glm_function = function(df) {
  df = 
    glm(solved_bin ~ victim_age + victim_race + victim_sex, data = df, family = binomial()) |> 
    broom::tidy() |> 
    filter(
      term == "victim_sexMale"
    ) |> 
    mutate(
      odds_ratio = exp(estimate),
      conf_upper_95 = exp(estimate + 1.96 * std.error),
      conf_lower_95 = exp(estimate - 1.96 * std.error)
    )
}

homicide_glm = 
  homicide_df |> 
  select(city_state, solved_bin, victim_sex, victim_age, victim_race) |>
  nest(data = solved_bin:victim_race) |> 
  mutate(
    glm_results = map(data, glm_function)
  ) |> 
  unnest(glm_results) |> 
  select(city_state, odds_ratio, conf_lower_95, conf_upper_95)

homicide_plot = 
  homicide_glm |> 
  ggplot(aes(x = reorder(city_state, odds_ratio), y = odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_lower_95, ymax = conf_upper_95)) 

print(homicide_plot)
```

## Problem 3
```{r}
bwt_df = 
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |>
  mutate(
    babysex = as.factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as.factor(frace),
    frace = fct_recode(
      frace, "white" = "1", "black" = "2", "asian" = "3", 
      "puerto rican" = "4", "other" = "8"),
    malform = as.logical(malform),
    mrace = as.factor(mrace),
    mrace = fct_recode(
      mrace, "white" = "1", "black" = "2", "asian" = "3", 
      "puerto rican" = "4"))
```
From a biological standpoint, I hypothesize that the following factors are more likely to underly birthweight: `babysex`, `bhead`, `blength`, `gaweeks`, `malform`, `smoken`. From a socioeconomical standpoint, I hypothesize that `frace` and `mrace` may also influence birthweight, because poverty, education, and employment tend to vary across different racial and ethnic groups.

#### Create my own model
I will first fit MLR with these variables as predictors.
```{r}
linear_model = lm(bwt ~ babysex + bhead + blength + gaweeks + malform + smoken + mrace + frace, data = bwt_df)
summary(linear_model)
```

The results suggest that all covariates are significant predictors of birthweight except for `frace` and `malform`. The adjusted R-squared for this model is 0.7049. I will proceed with refitting the model without these covariates.

```{r}
refitted_linear_model = lm(bwt ~ babysex + bhead + blength + gaweeks + smoken + mrace, data = bwt_df) 

summary(refitted_linear_model)
```
The refitted model now shows an adjusted R-squared of 0.7051. This suggests that the refitted model does a slightly better job at explaining the variation in birthweight compared to the original model. 

I am also interested in looking at interactions between different covariates. I will use `ggpairs` to look at multicollinearity between the continuous variables in my model. 

```{r}
ggpairs(bwt_df[, c("bhead", "blength", "smoken", "gaweeks")])
```

Results from ggpairs suggest that `blength` has a moderate linear relationship with `bhead` (corr = 0.630), `gaweeks` has a moderately weak relationship with `blength` (corr = 0.359) and `bhead` (corr = 0.378). I will fit another model that includes the three-way interaction between these covariates. 


```{r}
interaction_model = lm(bwt ~ babysex + bhead * blength * gaweeks + smoken + mrace, data = bwt_df)
summary(interaction_model)

AIC(interaction_model)
AIC(refitted_linear_model)
```

Results from interaction model suggest that all covariates are significant predictors of birthweight, and that the interactions between `gaweek`, `bhead` and `blength` are also significant (p-values are all lower than 0.05). The adjusted R-squared for this model is 0.7072, which suggests that this model is slightly better than the refitted linear model without interaction terms. 

I also investigated AIC values to evaluate goodness-of-fit of these two models, and found that the AIC value for the interaction model is slightly lower than that for the refitted linear model. Based on adjusted R-squared and AIC value, the interaction model will be my final model. 

The plot showing predictions against residuals is as follows:
```{r}
bwt_plot = 
  bwt_df |> 
  add_predictions(interaction_model) |> 
  add_residuals(interaction_model) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE,  color = "red") +
  labs(
    title = "Residuals vs. Predicted Birth Weight", x = "Predicted Birth Weight", y = "Residuals"
  )

print(bwt_plot)
```

The residual aginst prediction plot shows that there is an overall random scatter pattern, but the model does a better job at explaining variation in birth weight for birth weight values ranging from 2500 to 3500 grams, which is the typical birth weight for newborn babies. There are a few extreme data points around the range 500 to 1500, but this makes biological sense due to factors such as premature birth or maternal health conditions. 

#### Compare this model with other models. 

The first model, named `model1`, fits a MLR of `bwt` against `blength` and `gaweeks` as predictors. The second model, named `model2`, fits a MLR of `bwt` against `bhead`, `blength` and `babysex` including their three-way interaction. 
 
```{r}
model1 = lm(bwt ~ blength + gaweeks, data = bwt_df)
summary(model1)

model2 = lm(bwt ~ bhead * blength * babysex, data = bwt_df)
summary(model2)
```

Model 1 results show that both variables `blength` and `gaweeks` are significant predictors of birthweight, with both p-values being less than 2e-16. 

Model 2 results show that `bhead`, `blength` and `babysex` are all significant predictors of `bwt`. Based on interaction terms, those terms involving sex (`babysexfemale`) and the main predictors (`bhead`, `blength`) suggest that the effects of these variables on birthweight differ between male and female babies. Specifically, the three-way interaction (bhead:blength:babysexfemale) suggests that the effect of both head circumference and length at birth on birthweight is further modified by the baby's sex. This indicates that the relationship between `bwt` and main predictors `bhead` and `blength` are not the same for male and female babies. 

I will use `crossv_mc` to make comparisons between my model and the other models in terms of cross-validated prediction error. 

```{r}
cv_df = 
  crossv_mc(bwt_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_results_df = 
  cv_df |> 
  mutate(
    my_model = map(train, \(x) lm(bwt ~ babysex + bhead * blength * gaweeks + smoken + mrace, data = x)),
    model1 = map(train, \(x) lm(bwt ~ blength + gaweeks, data = x)), 
    model2 = map(train, \(x) lm(bwt ~ bhead * blength * babysex, data = x))
  ) |> 
  mutate(
    rmse_my_model = map2_dbl(my_model, test, rmse),
    rmse_model1 = map2_dbl(model1, test, rmse),
    rmse_model2 = map2_dbl(model2, test, rmse)
  ) |> 
  select(rmse_my_model, rmse_model1, rmse_model2)
 
head(cv_results_df) 
```

